Semi-parametric relative risk model for correlated binary data

### 2019-09-25 SPregress.tex
Experiment with fitting starting values

\section{Initial setup}
\begin{flushleft} \small\label{scrap1}\raggedright\small
\NWtarget{nuweb1a}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {1a}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@require(CorrBin)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1a}{1a}\NWlink{nuweb1b}{b}\NWlink{nuweb2}{, 2}\NWlink{nuweb3b}{, 3b}\NWlink{nuweb4}{, 4}\NWlink{nuweb5}{, 5}.

@@

\begin{flushleft} \small\label{scrap2}\raggedright\small
\NWtarget{nuweb1b}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {1b}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@# lambda-to-p weight matrix, rows:r, cols:s@\\
\mbox{}\verb@# (-1)^(s-r)*choose(n,r)*choose(n-r,s-r)@\\
\mbox{}\verb@@\\
\mbox{}\verb@weight_mat <- function(n){@\\
\mbox{}\verb@  s <- r <- 0:n@\\
\mbox{}\verb@  res <- outer(r, s, function(x,y)(-1)^(y-x)*choose(n,x)*choose(n-x,y-x))@\\
\mbox{}\verb@  res@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@p_from_lambda <- function(lambda.vec, n){@\\
\mbox{}\verb@  H <- weight_mat(n)@\\
\mbox{}\verb@  c(H %*% lambda.vec[1:(n+1)])@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@# p-to-lambda weight matrix, rows:k, cols:r@\\
\mbox{}\verb@# (choose(n-k, n-r)/choose(n,r)@\\
\mbox{}\verb@@\\
\mbox{}\verb@weight_mat2 <- function(n){@\\
\mbox{}\verb@  k <- r <- 0:n@\\
\mbox{}\verb@  res <- outer(k, r, function(x,y)choose(n-x, n-y)/choose(n,y))@\\
\mbox{}\verb@  res@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@lambda_from_p <- function(p.vec, n=length(p.vec)-1){@\\
\mbox{}\verb@  H <- weight_mat2(n)@\\
\mbox{}\verb@  c(H %*% p.vec)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1a}{1a}\NWlink{nuweb1b}{b}\NWlink{nuweb2}{, 2}\NWlink{nuweb3b}{, 3b}\NWlink{nuweb4}{, 4}\NWlink{nuweb5}{, 5}.

@@

\begin{flushleft} \small\label{scrap3}\raggedright\small
\NWtarget{nuweb2}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {2}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@#'@{\tt @}\verb@rdname sprr@\\
\mbox{}\verb@#'@{\tt @}\verb@param formula a one-sided formula of the form \code{cbind(r, s) ~ predictors} where \code{r} and \code{s} give the number of responses and non-responses within each cluster, respectively (so the cluster size is \code{r+s}), and \code{predictors} describes the covariates.@\\
\mbox{}\verb@#'@{\tt @}\verb@param data  an optional matrix or data frame containing the variables in the formula \code{formula}. By default the variables are taken from \code{environment(formula).}@\\
\mbox{}\verb@#'@{\tt @}\verb@param subset  an optional vector specifying a subset of observations to be used.@\\
\mbox{}\verb@#'@{\tt @}\verb@param weight  an optional vector specifying observation weights.@\\
\mbox{}\verb@#'@{\tt @}\verb@param link      a link function for the binomial distribution, the inverse of which models the covariate effects.@\\
\mbox{}\verb@#'@{\tt @}\verb@param mu1       an optional value between 0 and 1 giving the maximal predicted marginal probability. If set to NULL (default), the algorithm will try to estimate it from the data.@\\
\mbox{}\verb@#'@{\tt @}\verb@param start     an optional list with elements named \code{beta}, \code{q}, and/or \code{mu1} giving starting values for estimation. If none or only some are specified, starting values are selected automatically.@\\
\mbox{}\verb@#'@{\tt @}\verb@param control a list with parameters controlling the algorithm.@\\
\mbox{}\verb@#'@{\tt @}\verb@return an object of class \code{sprr} with the fitted model.@\\
\mbox{}\verb@#'@{\tt @}\verb@export@\\
\mbox{}\verb@#' @{\tt @}\verb@importFrom stats terms model.matrix@\\
\mbox{}\verb@@\\
\mbox{}\verb@#' Semi-parametric relative risk model@\\
\mbox{}\verb@sprr <- function(formula, data, subset, weights, link="cloglog", mu1=NULL, start=NULL, control=list(eps=0.001, maxit=100), ...){@\\
\mbox{}\verb@    fam <- binomial(link=link)@\\
\mbox{}\verb@    model_fun <- fam$linkinv@\\
\mbox{}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize \NWlink{nuweb3a}{3a}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Fit model}\nobreak\ {\footnotesize \NWlink{nuweb6}{6}}$\,\rangle$}\verb@@\\
\mbox{}\verb@@\\
\mbox{}\verb@    mt <- attr(mf, "terms")@\\
\mbox{}\verb@    res <- list(coefficients = beta_new, q=q_new, niter = iter, loglik=logl,@\\
\mbox{}\verb@                link = link, call = mc, terms = mt,@\\
\mbox{}\verb@                xlevels = .getXlevels(mt, mf),@\\
\mbox{}\verb@                data_object=list(model_matrix=mm, resp=Y, n=rowSums(Y), weights=weights),@\\
\mbox{}\verb@                model_fun=model_fun)@\\
\mbox{}\verb@    class(res) <- "sprr"@\\
\mbox{}\verb@    res@\\
\mbox{}\verb@@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1a}{1a}\NWlink{nuweb1b}{b}\NWlink{nuweb2}{, 2}\NWlink{nuweb3b}{, 3b}\NWlink{nuweb4}{, 4}\NWlink{nuweb5}{, 5}.

\item{}
\end{list}
\vspace{4ex}
\end{flushleft}
\begin{flushleft} \small\label{scrap4}\raggedright\small
\NWtarget{nuweb3a}{} $\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize {3a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@   if (missing(formula) || (length(formula) != 3L))@\\
\mbox{}\verb@        stop("'formula' missing or incorrect")@\\
\mbox{}\verb@   if (missing(data))@\\
\mbox{}\verb@        data <- environment(formula)@\\
\mbox{}\verb@    mc <- match.call(expand.dots = FALSE)@\\
\mbox{}\verb@    m <- match(c("formula", "data", "subset", "weights"), names(mc), 0L)@\\
\mbox{}\verb@    m <- mc[c(1L,m)]@\\
\mbox{}\verb@    if (is.matrix(eval(m$data, parent.frame())))@\\
\mbox{}\verb@        m$data <- as.data.frame(data)@\\
\mbox{}\verb@    m[[1L]] <- quote(stats::model.frame)@\\
\mbox{}\verb@    mf <- eval(m, parent.frame())@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # extract and check response variable@\\
\mbox{}\verb@    Y <- model.response(mf)@\\
\mbox{}\verb@    if (ncol(Y) != 2) stop("The response variable should be a 2-column matrix")@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # create model matrix@\\
\mbox{}\verb@    mm <- model.matrix(formula, data=mf)@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # extract weights@\\
\mbox{}\verb@    weights <- as.vector(model.weights(mf))@\\
\mbox{}\verb@    if (!is.null(weights) && !is.numeric(weights))@\\
\mbox{}\verb@        stop("'weights' must be a numeric vector")@\\
\mbox{}\verb@    if (!is.null(weights) && any(weights < 0))@\\
\mbox{}\verb@        stop("negative weights not allowed")@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb2}{2}.

@@

\begin{flushleft} \small\label{scrap5}\raggedright\small
\NWtarget{nuweb3b}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {3b}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@pred_lp <- function(beta, data_object){@\\
\mbox{}\verb@    lp <- data_object$model_matrix %*% beta@\\
\mbox{}\verb@    c(lp)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_theta <- function(beta, data_object, model_fun){@\\
\mbox{}\verb@    lp <- pred_lp(beta, data_object)@\\
\mbox{}\verb@    theta <- model_fun(lp)@\\
\mbox{}\verb@    theta@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@pred_lambda <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@    lp <- data_object$model_matrix %*% beta@\\
\mbox{}\verb@    theta <- model_fun(lp)@\\
\mbox{}\verb@    N <- length(q)-1@\\
\mbox{}\verb@    lambda_N <- lambda_from_p(q)@\\
\mbox{}\verb@    th <- sapply(0:N, function(k)theta^k)@\\
\mbox{}\verb@    lambda <- apply(th, 1, function(t)t * lambda_N)@\\
\mbox{}\verb@    rownames(lambda) <- 0:N@\\
\mbox{}\verb@    t(lambda)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_pvec <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@   ll <- pred_lambda(beta, q, data_object, model_fun)@\\
\mbox{}\verb@   cs <- data_object$n@\\
\mbox{}\verb@   pp <- lapply(seq_along(cs), @\\
\mbox{}\verb@                function(i){res <- p_from_lambda(ll[i,], n=cs[i]);@\\
\mbox{}\verb@                            names(res) <- 0:cs[i];@\\
\mbox{}\verb@                            res})@\\
\mbox{}\verb@   pp@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_p <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@   pvec <- pred_pvec(beta, q, data_object, model_fun)@\\
\mbox{}\verb@   rvec <- data_object$resp[,1]@\\
\mbox{}\verb@   pp <- sapply(seq_along(rvec), function(i)pvec[[i]][rvec[i]+1])@\\
\mbox{}\verb@   pp@\\
\mbox{}\verb@}@\\
\mbox{}\verb@loglik <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@    p <- pred_p(beta=beta, q=q, data_object = data_object, model_fun=model_fun)@\\
\mbox{}\verb@    w <- data_object$weights@\\
\mbox{}\verb@    if (is.null(w)) ll <- sum(log(p))@\\
\mbox{}\verb@    else ll <- sum(ifelse(w==0, 0, w*log(p)))@\\
\mbox{}\verb@    ll@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1a}{1a}\NWlink{nuweb1b}{b}\NWlink{nuweb2}{, 2}\NWlink{nuweb3b}{, 3b}\NWlink{nuweb4}{, 4}\NWlink{nuweb5}{, 5}.

@@

\begin{flushleft} \small\label{scrap6}\raggedright\small
\NWtarget{nuweb4}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {4}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@# based on print.lm@\\
\mbox{}\verb@print.sprr <- function(x, digits = max(3L, getOption("digits") - 3L),...){@\\
\mbox{}\verb@  cat("\nA semi-parametric relative risk regression model fit\n")@\\
\mbox{}\verb@  cat("\nCall:\n", paste(deparse(x$call), sep = "\n", collapse = "\n"),@\\
\mbox{}\verb@        "\n\n", sep = "")@\\
\mbox{}\verb@  if (length(coef(x))) {@\\
\mbox{}\verb@    cat("Coefficients:\n")@\\
\mbox{}\verb@    print.default(format(x$coefficients, digits = digits),@\\
\mbox{}\verb@            print.gap = 2, quote = FALSE)@\\
\mbox{}\verb@    }@\\
\mbox{}\verb@    else cat("No coefficients\n")@\\
\mbox{}\verb@@\\
\mbox{}\verb@    if (length(x$q)){@\\
\mbox{}\verb@        cat("Baseline joint probabilities (mu):\n")     @\\
\mbox{}\verb@        N <- length(x$q)-1@\\
\mbox{}\verb@        ll <- c(lambda_from_p(x$q))@\\
\mbox{}\verb@        names(ll) <- 0:N@\\
\mbox{}\verb@        print.default(format(ll, digits = digits),@\\
\mbox{}\verb@            print.gap = 2, quote = FALSE)@\\
\mbox{}\verb@    }@\\
\mbox{}\verb@    else cat("No baseline joint probabilities\n")@\\
\mbox{}\verb@@\\
\mbox{}\verb@   cat("Log-likelihood: ", format(x$loglik, digits=digits), "\n")@\\
\mbox{}\verb@   invisible(x)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1a}{1a}\NWlink{nuweb1b}{b}\NWlink{nuweb2}{, 2}\NWlink{nuweb3b}{, 3b}\NWlink{nuweb4}{, 4}\NWlink{nuweb5}{, 5}.

@@

\begin{flushleft} \small\label{scrap9}\raggedright\small
\NWtarget{nuweb8a}{} $\langle\,${\itshape Setup for E-step}\nobreak\ {\footnotesize {8a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@    # replace each cluster with N(N-1)/2 clusters of size N@\\
\mbox{}\verb@    new <- cbind(s=rep(0:N, each=N+1), y=rep(0:N, times=N+1))@\\
\mbox{}\verb@    new <- new[new[,"s"] <= new[,"y"],]@\\
\mbox{}\verb@@\\
\mbox{}\verb@    rep_idx <- rep(1:nrow(Y), each=nrow(new))@\\
\mbox{}\verb@    Y2 <- cbind(1:nrow(Y), Y)[rep_idx,]@\\
\mbox{}\verb@    colnames(Y2) <- c("i", "resp","nonresp")@\\
\mbox{}\verb@    mm2 <- mm[rep_idx,,drop=FALSE]@\\
\mbox{}\verb@@\\
\mbox{}\verb@    rep_idx2 <- rep(1:nrow(new), times=nrow(Y))@\\
\mbox{}\verb@    Ycomb <- cbind(Y2, new[rep_idx2,])@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb6}{6}.

>>>>

\section{Initial setup}
\begin{flushleft} \small\label{scrap1}\raggedright\small
\NWtarget{nuweb?}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {?}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@require(CorrBin)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb?}{?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}.

@@

\begin{flushleft} \small\label{scrap2}\raggedright\small
\NWtarget{nuweb?}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {?}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@# lambda-to-p weight matrix, rows:r, cols:s@\\
\mbox{}\verb@# (-1)^(s-r)*choose(n,r)*choose(n-r,s-r)@\\
\mbox{}\verb@@\\
\mbox{}\verb@weight_mat <- function(n){@\\
\mbox{}\verb@  s <- r <- 0:n@\\
\mbox{}\verb@  res <- outer(r, s, function(x,y)(-1)^(y-x)*choose(n,x)*choose(n-x,y-x))@\\
\mbox{}\verb@  res@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@p_from_lambda <- function(lambda.vec, n){@\\
\mbox{}\verb@  H <- weight_mat(n)@\\
\mbox{}\verb@  c(H %*% lambda.vec[1:(n+1)])@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@# p-to-lambda weight matrix, rows:k, cols:r@\\
\mbox{}\verb@# (choose(n-k, n-r)/choose(n,r)@\\
\mbox{}\verb@@\\
\mbox{}\verb@weight_mat2 <- function(n){@\\
\mbox{}\verb@  k <- r <- 0:n@\\
\mbox{}\verb@  res <- outer(k, r, function(x,y)choose(n-x, n-y)/choose(n,y))@\\
\mbox{}\verb@  res@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@lambda_from_p <- function(p.vec, n=length(p.vec)-1){@\\
\mbox{}\verb@  H <- weight_mat2(n)@\\
\mbox{}\verb@  c(H %*% p.vec)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb?}{?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}.

@@

\begin{flushleft} \small\label{scrap3}\raggedright\small
\NWtarget{nuweb?}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {?}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@#'@{\tt @}\verb@rdname sprr@\\
\mbox{}\verb@#'@{\tt @}\verb@param formula a one-sided formula of the form \code{cbind(r, s) ~ predictors} where \code{r} and \code{s} give the number of responses and non-responses within each cluster, respectively (so the cluster size is \code{r+s}), and \code{predictors} describes the covariates.@\\
\mbox{}\verb@#'@{\tt @}\verb@param data  an optional matrix or data frame containing the variables in the formula \code{formula}. By default the variables are taken from \code{environment(formula).}@\\
\mbox{}\verb@#'@{\tt @}\verb@param subset  an optional vector specifying a subset of observations to be used.@\\
\mbox{}\verb@#'@{\tt @}\verb@param weight  an optional vector specifying observation weights.@\\
\mbox{}\verb@#'@{\tt @}\verb@param link      a link function for the binomial distribution, the inverse of which models the covariate effects.@\\
\mbox{}\verb@#'@{\tt @}\verb@param mu1       an optional value between 0 and 1 giving the maximal predicted marginal probability. If set to NULL (default), the algorithm will try to estimate it from the data.@\\
\mbox{}\verb@#'@{\tt @}\verb@param start     an optional list with elements named \code{beta}, \code{q}, and/or \code{mu1} giving starting values for estimation. If none or only some are specified, starting values are selected automatically.@\\
\mbox{}\verb@#'@{\tt @}\verb@param control a list with parameters controlling the algorithm.@\\
\mbox{}\verb@#'@{\tt @}\verb@return an object of class \code{sprr} with the fitted model.@\\
\mbox{}\verb@#'@{\tt @}\verb@export@\\
\mbox{}\verb@#' @{\tt @}\verb@importFrom stats terms model.matrix@\\
\mbox{}\verb@@\\
\mbox{}\verb@#' Semi-parametric relative risk model@\\
\mbox{}\verb@sprr <- function(formula, data, subset, weights, link="cloglog", mu1=NULL, start=NULL, control=list(eps=0.001, maxit=100), ...){@\\
\mbox{}\verb@    fam <- binomial(link=link)@\\
\mbox{}\verb@    model_fun <- fam$linkinv@\\
\mbox{}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize \NWlink{nuweb?}{?}}$\,\rangle$}\verb@@\\
\mbox{}\verb@    @\hbox{$\langle\,${\itshape Fit model}\nobreak\ {\footnotesize \NWlink{nuweb?}{?}}$\,\rangle$}\verb@@\\
\mbox{}\verb@@\\
\mbox{}\verb@    mt <- attr(mf, "terms")@\\
\mbox{}\verb@    res <- list(coefficients = beta_new, q=q_new, niter = iter, loglik=logl,@\\
\mbox{}\verb@                link = link, call = mc, terms = mt,@\\
\mbox{}\verb@                xlevels = .getXlevels(mt, mf),@\\
\mbox{}\verb@                data_object=list(model_matrix=mm, resp=Y, n=rowSums(Y), weights=weights),@\\
\mbox{}\verb@                model_fun=model_fun)@\\
\mbox{}\verb@    class(res) <- "sprr"@\\
\mbox{}\verb@    res@\\
\mbox{}\verb@@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb?}{?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}.

\item{}
\end{list}
\vspace{4ex}
\end{flushleft}
\begin{flushleft} \small\label{scrap4}\raggedright\small
\NWtarget{nuweb?}{} $\langle\,${\itshape Create model matrix from formula and data}\nobreak\ {\footnotesize {?}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@   if (missing(formula) || (length(formula) != 3L))@\\
\mbox{}\verb@        stop("'formula' missing or incorrect")@\\
\mbox{}\verb@   if (missing(data))@\\
\mbox{}\verb@        data <- environment(formula)@\\
\mbox{}\verb@    mc <- match.call(expand.dots = FALSE)@\\
\mbox{}\verb@    m <- match(c("formula", "data", "subset", "weights"), names(mc), 0L)@\\
\mbox{}\verb@    m <- mc[c(1L,m)]@\\
\mbox{}\verb@    if (is.matrix(eval(m$data, parent.frame())))@\\
\mbox{}\verb@        m$data <- as.data.frame(data)@\\
\mbox{}\verb@    m[[1L]] <- quote(stats::model.frame)@\\
\mbox{}\verb@    mf <- eval(m, parent.frame())@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # extract and check response variable@\\
\mbox{}\verb@    Y <- model.response(mf)@\\
\mbox{}\verb@    if (ncol(Y) != 2) stop("The response variable should be a 2-column matrix")@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # create model matrix@\\
\mbox{}\verb@    mm <- model.matrix(formula, data=mf)@\\
\mbox{}\verb@@\\
\mbox{}\verb@    # extract weights@\\
\mbox{}\verb@    weights <- as.vector(model.weights(mf))@\\
\mbox{}\verb@    if (!is.null(weights) && !is.numeric(weights))@\\
\mbox{}\verb@        stop("'weights' must be a numeric vector")@\\
\mbox{}\verb@    if (!is.null(weights) && any(weights < 0))@\\
\mbox{}\verb@        stop("negative weights not allowed")@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb?}{?}.

@@

\begin{flushleft} \small\label{scrap5}\raggedright\small
\NWtarget{nuweb?}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {?}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@pred_lp <- function(beta, data_object){@\\
\mbox{}\verb@    lp <- data_object$model_matrix %*% beta@\\
\mbox{}\verb@    c(lp)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_theta <- function(beta, data_object, model_fun){@\\
\mbox{}\verb@    lp <- pred_lp(beta, data_object)@\\
\mbox{}\verb@    theta <- model_fun(lp)@\\
\mbox{}\verb@    theta@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@pred_lambda <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@    lp <- data_object$model_matrix %*% beta@\\
\mbox{}\verb@    theta <- model_fun(lp)@\\
\mbox{}\verb@    N <- length(q)-1@\\
\mbox{}\verb@    lambda_N <- lambda_from_p(q)@\\
\mbox{}\verb@    th <- sapply(0:N, function(k)theta^k)@\\
\mbox{}\verb@    lambda <- apply(th, 1, function(t)t * lambda_N)@\\
\mbox{}\verb@    rownames(lambda) <- 0:N@\\
\mbox{}\verb@    t(lambda)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_pvec <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@   ll <- pred_lambda(beta, q, data_object, model_fun)@\\
\mbox{}\verb@   cs <- data_object$n@\\
\mbox{}\verb@   pp <- lapply(seq_along(cs), @\\
\mbox{}\verb@                function(i){res <- p_from_lambda(ll[i,], n=cs[i]);@\\
\mbox{}\verb@                            names(res) <- 0:cs[i];@\\
\mbox{}\verb@                            res})@\\
\mbox{}\verb@   pp@\\
\mbox{}\verb@}@\\
\mbox{}\verb@pred_p <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@   pvec <- pred_pvec(beta, q, data_object, model_fun)@\\
\mbox{}\verb@   rvec <- data_object$resp[,1]@\\
\mbox{}\verb@   pp <- sapply(seq_along(rvec), function(i)pvec[[i]][rvec[i]+1])@\\
\mbox{}\verb@   pp@\\
\mbox{}\verb@}@\\
\mbox{}\verb@loglik <- function(beta, q, data_object, model_fun){@\\
\mbox{}\verb@    p <- pred_p(beta=beta, q=q, data_object = data_object, model_fun=model_fun)@\\
\mbox{}\verb@    w <- data_object$weights@\\
\mbox{}\verb@    if (is.null(w)) ll <- sum(log(p))@\\
\mbox{}\verb@    else ll <- sum(ifelse(w==0, 0, w*log(p)))@\\
\mbox{}\verb@    ll@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb?}{?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}.

@@

\begin{flushleft} \small\label{scrap6}\raggedright\small
\NWtarget{nuweb?}{} \verb@"R/SPreg.R"@\nobreak\ {\footnotesize {?}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@# based on print.lm@\\
\mbox{}\verb@print.sprr <- function(x, digits = max(3L, getOption("digits") - 3L),...){@\\
\mbox{}\verb@  cat("\nA semi-parametric relative risk regression model fit\n")@\\
\mbox{}\verb@  cat("\nCall:\n", paste(deparse(x$call), sep = "\n", collapse = "\n"),@\\
\mbox{}\verb@        "\n\n", sep = "")@\\
\mbox{}\verb@  if (length(coef(x))) {@\\
\mbox{}\verb@    cat("Coefficients:\n")@\\
\mbox{}\verb@    print.default(format(x$coefficients, digits = digits),@\\
\mbox{}\verb@            print.gap = 2, quote = FALSE)@\\
\mbox{}\verb@    }@\\
\mbox{}\verb@    else cat("No coefficients\n")@\\
\mbox{}\verb@@\\
\mbox{}\verb@    if (length(x$q)){@\\
\mbox{}\verb@        cat("Baseline joint probabilities (mu):\n")     @\\
\mbox{}\verb@        N <- length(x$q)-1@\\
\mbox{}\verb@        ll <- c(lambda_from_p(x$q))@\\
\mbox{}\verb@        names(ll) <- 0:N@\\
\mbox{}\verb@        print.default(format(ll, digits = digits),@\\
\mbox{}\verb@            print.gap = 2, quote = FALSE)@\\
\mbox{}\verb@    }@\\
\mbox{}\verb@    else cat("No baseline joint probabilities\n")@\\
\mbox{}\verb@@\\
\mbox{}\verb@   cat("Log-likelihood: ", format(x$loglik, digits=digits), "\n")@\\
\mbox{}\verb@   invisible(x)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb?}{?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}\NWlink{nuweb?}{, ?}.

@@

\begin{flushleft} \small\label{scrap9}\raggedright\small
\NWtarget{nuweb?}{} $\langle\,${\itshape Setup for E-step}\nobreak\ {\footnotesize {?}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@    # replace each cluster with N(N-1)/2 clusters of size N@\\
\mbox{}\verb@    new <- cbind(s=rep(0:N, each=N+1), y=rep(0:N, times=N+1))@\\
\mbox{}\verb@    new <- new[new[,"s"] <= new[,"y"],]@\\
\mbox{}\verb@@\\
\mbox{}\verb@    rep_idx <- rep(1:nrow(Y), each=nrow(new))@\\
\mbox{}\verb@    Y2 <- cbind(1:nrow(Y), Y)[rep_idx,]@\\
\mbox{}\verb@    colnames(Y2) <- c("i", "resp","nonresp")@\\
\mbox{}\verb@    mm2 <- mm[rep_idx,,drop=FALSE]@\\
\mbox{}\verb@@\\
\mbox{}\verb@    rep_idx2 <- rep(1:nrow(new), times=nrow(Y))@\\
\mbox{}\verb@    Ycomb <- cbind(Y2, new[rep_idx2,])@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb?}{?}.
